```{r}
#| warning: false
#| message: false
source("R/Vars.R")
```

# Placing Analysis

This page seeks to see if there are underlying trends based one what placing a competitor took in both the Absolute class and their own weight class.

## Weight Class Breakdown

When we look at each individual placing, 1 through 3, in the absolute and compare it to the weight weight class of the competitor a few insights appear.

1.  We see the general trend of Medium Heavy to Super Heavy being the weight classes that perform the best that we saw in our overview still holds even at the individual placing level.
2.  More intrestingly we see taking 2nd and 3rd in the absolute have very similair breakdowns of weight classes, but for taking 1st we see a much higher proportion of competitors fall in the Heavy and particularly the Super Heavy classes, with the Medium Heavy weight class having a significant drop off. The Super Heavy and Heavy weight classes by themselves make up 50% of the 1st place winners.\

```{r}
#| label: PlacingAnalysis_Plot_Absolute
read_Plot("Placings_Absolute.rds")
```

## Comparison of Weight Class Placing and Absolute Placing

When looking at the relationship between how well a competitor placed in their weight class in comparison to the Absolute class, unsurprisingly we see that the better a competitor does in their weight class the more likely they are to do well in the absolute class. Although it is still striking to see that nearly %80 of 1st place winners of the absolute took 1st int their weight class. In short taking 1st in your weight class is a very strong signal you will perform well in the absolute.

```{r}
#| label: PlacingAnalysis_Plot_WeightsClass
read_Plot("Placings_Weight_Class.rds")
```

When we look at the overall distribution of Competitors that made the podium in the absolute class, we see the trend of 1st place winners from the weight class performing well in the absolute holds. Out of all competitors who placed in the absolute at all the vast majority were those who had placed 1st in their weight class division.

```{r}
#| label: PlacingAnalysis_Plot_Distribution
read_Plot("Placings_Distribution.rds")
```
